---
title: "RWorksheet_Junsay#3b"
author: "Syndric James Z. Junsay"
date: "2025-10-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 1. Create a data frame using the table below.
 a. Write the codes.
 
```{r}
Respondents <- 1:20
Sex <- c(2,2,1,2,2,2,2,2,2,1,1,2,2,2,2,2,2,1,1,2)
Fathers_Occupation <- c(1,3,3,3,1,2,3,1,1,1,3,2,1,3,3,1,3,1,2,1)
Persons_at_Home <- c(5,7,3,8,1,9,6,7,8,4,7,5,4,7,8,8,3,11,5,6)
Siblings_at_School <- c(6,4,4,1,4,1,5,3,1,2,3,2,5,3,2,1,2,5,3,2)
Types_of_Houses <- c(1,2,3,1,1,3,3,1,3,2,2,3,2,3,3,3,3,3,3,2)

# Combine into a data frame
survey_data <- data.frame(Respondents, Sex, Fathers_Occupation, Persons_at_Home, 
                          Siblings_at_School, Types_of_Houses)

# Display the data
print(survey_data)
```
b.Describe the data. Get the structure or the summary of the data

```{r}
str(survey_data)
summary(survey_data)
```
c. Is the mean number of siblings attending is 5?
```{r}
mean(Siblings_at_School)
```
no, the mean number of siblings attending is not 5.

d. Extract the 1st two rows and then all the columns using the subsetting functions.
Write the codes and its output.

```{r}
first_two <- survey_data[1:2, ]
print(first_two)
```
e. Extract 3rd and 5th row with 2nd and 4th column. Write the codes and its result.
```{r}
third_fourth <- survey_data[c(3, 5), c(2,4)]
print(third_fourth)
```
f. Select the variable types of houses then store the vector that results as types_houses.
Write the codes.
```{r}
types_houses <- survey_data$Types_of_Houses
print(types_houses)
```
g. Select only all Males respondent that their father occupation was farmer. Write the
codes and its output.
```{r}
male_farmers <- subset(survey_data, Sex == 1 & Fathers_Occupation == 1)
print(male_farmers)
```
h. Select only all females respondent that have greater than or equal to 5 number of
siblings attending school. Write the codes and its outputs.
```{r}
female_many_siblings <- subset(survey_data, Sex == 2 & Siblings_at_School >= 5)
print(female_many_siblings)
```
## 2. Write a R program to create an empty data frame. Using the following codes:

a. Describe the results.
```{r}
df = data.frame(Ints=integer(),Doubles=double(), Characters=character(), Logicals=logical(), Factors=factor(), stringsAsFactors=FALSE)
print("Structure of the empty dataframe:")
print(str(df))
```
3. Create a .csv file of this. Save it as HouseholdData.csv
```{r}
Respondents <- 1:10
Sex <- c("Male", "Female", "Female", "Male", "Male", 
         "Female", "Female", "Male", "Female", "Male")
Fathers_Occupation <- c(1, 2, 3, 3, 1, 2, 2, 3, 1, 3)
Persons_at_Home <- c(5, 7, 3, 8, 6, 4, 4, 2, 11, 6)
Siblings_at_School <- c(2, 3, 0, 5, 2, 3, 1, 2, 6, 2)
Types_of_Houses <- c("Wood", "Congrete", "Congrete", "Wood", 
                     "Semi-congrete", "Semi-congrete", "Wood", 
                     "Semi-congrete", "Semi-congrete", "Congrete")

# Combine into a data frame
HouseholdData <- data.frame(Respondents, Sex, Fathers_Occupation,
                            Persons_at_Home, Siblings_at_School,
                            Types_of_Houses)

# View the data frame
print(HouseholdData)

write.csv(HouseholdData, "HouseholdData.csv", row.names = FALSE)
```

a. Import the csv file into the R environment. Write the codes.
```{r}
HouseholdData <- read.csv("HouseholdData.csv")
print(HouseholdData)
```
b. Convert the Sex into factor using factor() function and change it into integer.[Legend:
Male = 1 and Female = 2]. Write the R codes and its output.
```{r}
HouseholdData$Sex <- factor(HouseholdData$Sex,
                            levels = c("Male", "Female"),
                            labels = c(1, 2))
print(HouseholdData)
str(HouseholdData)
```
c. Convert the Type of Houses into factor and change it into integer. [Legend: Wood
= 1; Congrete = 2; Semi-Congrete = 3]. Write the R codes and its output.
```{r}
HouseholdData$Types_of_Houses <- factor(HouseholdData$Types_of_Houses,
                                        levels = c("Wood", "Congrete", "Semi-congrete"),
                                        labels = c(1, 2, 3))
print(HouseholdData)
str(HouseholdData)
```

d. On father’s occupation, factor it as Farmer = 1; Driver = 2; and Others = 3. What
is the R code and its output?
```{r}
HouseholdData$Fathers_Occupation <- factor(HouseholdData$Fathers_Occupation,
                                           levels = c(1, 2, 3),
                                           labels = c("Farmer", "Driver", "Others"))
print(HouseholdData)
str(HouseholdData)
```
e. Select only all females respondent that has a father whose occupation is driver. Write
the codes and its output.
```{r}
female_driver <- subset(HouseholdData, Sex == 2 & Fathers_Occupation == "Driver")
print(female_driver)
```
f. Select the respondents that have greater than or equal to 5 number of siblings attending
school. Write the codes and its output.

```{r}
five_or_more_siblings <- subset(HouseholdData, Siblings_at_School >= 5)
print(five_or_more_siblings)
```
##4. Interpret the graph.
```{r}
library(ggplot2)

sentiments <- data.frame(
  Date = rep(c("July 14, 2020", "July 15, 2020", "July 17, 2020",
               "July 18, 2020", "July 20, 2020", "July 21, 2020"), each = 3),
  Sentiment = rep(c("Negative", "Neutral", "Positive"), times = 6),
  Count = c(2500, 1500, 1800,
            4200, 2800, 3200,
            3300, 2100, 2500,
            3200, 2000, 2600,
            2300, 1500, 2000,
            4100, 2700, 3400)
)


ggplot(sentiments, aes(x = interaction(Date, Sentiment), y = Count, fill = Sentiment)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Sentiments Of Tweets Per Day",
       x = "Day of Date / Sentiment",
       y = "Count of Tweets") +
  scale_fill_manual(values = c("red", "orange", "blue")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
## The graph shows that tweet sentiments were mostly negative at the beginning of the period, gradually shifted to more neutral tones midweek, and became increasingly positive toward the end.









