---
title: "RWorksheet_Junsay#4a.Rmd"
author: "Syndric James Z. Junsay"
date: "2025-11-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##1. The table below shows the data about shoe size and height. Create a data frame.

```{r}
data <- data.frame(
  ShoeSize = c(6.5, 9.0, 8.5, 8.5, 10.5, 9.0, 9.5, 7.5, 7.5, 10.5,
               8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5,
               8.5, 10.5, 11.0, 9.0, 13.0),
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 64.0, 72.0,
             67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 64.0, 64.0,
             67.0, 69.0, 70.0, 69.0, 70.0),
  Gender = c("F", "F", "F", "M", "M", "F", "F", "F", "M", "M",
             "F", "M", "M", "M", "M", "F", "F", "M", "F", "M",
             "M", "M", "M", "M", "M")
)
print(data)
```

a. Describe the data.

The dataset includes 25 people with their shoe sizes, heights, and genders. Males are generally taller and have larger shoe sizes than females, and there is a clear positive relationship between shoe size and height.

b. Create a subset by males and females with their corresponding shoe size and height. What its result? Show the R scripts.

```{r}
# Subset for males
male_data <- subset(data, Gender == "M", select = c(ShoeSize, Height))

# Subset for females
female_data <- subset(data, Gender == "F", select = c(ShoeSize, Height))

male_data
female_data
```
c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.

```{r}
# Mean of shoe size
mean_shoesize <- mean(data$ShoeSize)

# Mean of height
mean_height <- mean(data$Height)

mean_shoesize
mean_height
```
d. Is there a relationship between shoe size and height? Why?

Yes there is a relationship between shoe size and height.
As the shoe size increases, the height also tends to increase. This pattern suggests that taller individuals usually have larger feet, which makes sense biologically because body proportions are generally related. In other words, people with bigger bodies (greater height) often require larger shoe sizes to support their frame.

##2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out the factor levels below the actual values.

```{r}
months <- c("March", "April", "January", "November", "January",
             "September", "October", "September", "November", "August",
             "January", "November", "November", "February", "May", "August",
             "July", "December", "August", "August", "September", "November",
             "February", "April")


factor_months_vector <- factor(months)


print(factor_months_vector)
```

##3. Then check the summary() of the months_vector and factor_months_vector. | Inter-pret the results of both vectors. Are they both equally useful in this case?

```{r}
months_vector <- c("March", "April", "January", "November", "January",
                   "September", "October", "September", "November", "August",
                   "January", "November", "November", "February", "May", "August",
                   "July", "December", "August", "August", "September", "November",
                   "February", "April")


factor_months_vector <- factor(months_vector)

summary(months_vector)
summary(factor_months_vector)
```
##4. Create a vector and factor for the table below.

```{r}
direction <- c("East", "West", "West", "West", "West", 
               "North", "North", "North")

factor_data <- factor(direction)

new_order_data <- factor(factor_data, levels = c("East", "West", "North"))

print(new_order_data)
```
##5. Enter the data below in Excel with file name = import_march.csv

```{r}
getwd()
```
a. Import the excel file into the
```{r}
import_data <- read.table("import_march.csv", 
                          sep = ",", 
                          header = TRUE)

```
b. View the dataset. Write the R scripts and its result.

```{r}
print(data)
```

##6. Full Search
a. Create an R Program that allows the User to randomly select numbers from 1 to 50. Then display the chosen number. If the number is beyond the range of the selected choice, it will have to display a string “The number selected is beyond the range of 1 to 50”. If
number 20 is inputted by the User, it will have to display “TRUE”, otherwise display the input number.

```{r}
# Function that performs exhaustive (one-by-one) search in the set 1:50
exhaustive_search <- function(target) {
  domain <- 1:50
  found <- FALSE
  
  # Check each element one-by-one (exhaustive)
  for (i in seq_along(domain)) {
    if (domain[i] == target) {
      found <- TRUE
      break
    }
  }
  
  return(found)
}

# Main interaction: choose mode
cat("Choose mode:\n1 - Randomly select a number (1..50)\n2 - Manually input a number\n")
mode <- readline(prompt = "Enter 1 or 2: ")

if (mode == "1") {
  # Random pick from 1..50
  chosen <- sample(1:50, 1)
  cat("Randomly chosen number:", chosen, "\n")
} else {
  # Manual input
  user_input <- readline(prompt = "Type a number: ")
  # try to coerce to numeric; NA if non-numeric
  chosen <- suppressWarnings(as.numeric(user_input))
  if (is.na(chosen)) {
    cat("Invalid input: please type a numeric value.\n")
    quit(save = "no", status = 1)
  }
  cat("You typed:", chosen, "\n")
}

# Validate range
if (chosen < 1 || chosen > 50) {
  cat("The number selected is beyond the range of 1 to 50\n")
} else {
  # Do exhaustive search to demonstrate checking all elements
  found <- exhaustive_search(as.integer(chosen))
  
  if (found) {
    # Special rule: if number is 20 print TRUE, otherwise print the number
    if (as.integer(chosen) == 20L) {
      cat("TRUE\n")
    } else {
      cat(chosen, "\n")
    }
  } else {
    # This branch should not happen for chosen in 1..50, but kept for completeness
    cat("Number not found in the set 1..50\n")
  }
}
```

##7. Change
a. Write a function that prints the minimum number of bills that must be paid, given the
price of the snack.

```{r}
min_bills <- function(price) {
  # Available bills in descending order
  bills <- c(1000, 500, 200, 100, 50)
  
  remaining <- price
  count <- 0
  
  # Greedy approach: always use the largest bill possible
  for (b in bills) {
    # Determine how many bills of this type can be used
    num_bills <- remaining %/% b  
    count <- count + num_bills
    remaining <- remaining %% b  
  }
  
  return(count)
}
```

##8. The following is each student’s math score for one semester. Based on this, answer the following questions.

a. Create a dataframe from the above table. Write the R codes and its output.

```{r}
students <- data.frame(
  Name   = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)

# Display the dataframe
students
```

b. Without using the rowMean function, output the average score of students whose average
math score over 90 points during the semester. write R code and its output.

Example Output: Annie’s average grade this semester is 88.75.

```{r}
students <- data.frame(
  Name   = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)

# Compute averages manually (NO rowMeans)
for (i in 1:nrow(students)) {
  grades <- c(students$Grade1[i], students$Grade2[i],
              students$Grade3[i], students$Grade4[i])
  
  avg <- sum(grades) / length(grades)
  
  # Print only if average > 90
  if (avg > 90) {
    cat(students$Name[i], "'s average grade this semester is ", avg, ".\n", sep = "")
  }
}
```

c. Without using the mean function, output as follows for the tests in which the average score was less than 80 out of 4 tests.

Example output: The nth test was difficult.

```{r}
students <- data.frame(
  Name   = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)

# Loop through each test (column 2 to 5)
for (test in 2:5) {
  
  # Extract the scores of the current test
  scores <- students[, test]
  
  # Compute the average manually
  avg <- sum(scores) / length(scores)
  
  # If the average of the test is below 80, print message
  if (avg < 80) {
    test_number <- test - 1   # Because column 2 = Test 1
    cat("The", test_number, "test was difficult.\n")
  }
}

```

d. Without using the max function, output as follows for students whose highest score for a semester exceeds 90 points.

Example Output: Annie’s highest grade this semester is 95.

```{r}
students <- data.frame(
  Name   = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)

# Loop through each student
for (i in 1:nrow(students)) {
  
  # Collect the four grades
  grades <- c(students$Grade1[i], students$Grade2[i],
              students$Grade3[i], students$Grade4[i])
  
  # Find highest value manually (no max function)
  highest <- grades[1]
  for (g in grades) {
    if (g > highest) {
      highest <- g
    }
  }
  
  # Print only if highest score > 90
  if (highest > 90) {
    cat(students$Name[i], "'s highest grade this semester is ", highest, ".\n", sep = "")
  }
}
```


