---
title: "RWorksheet_Junsay#4b.Rmd"
author: "Syndric James Z. Junsay"
date: "2025-11-24"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute
```{r}
# Define the vector
vectorA <- 1:5
matrix_result <- matrix(0, nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
    matrix_result[i, j] <- abs(i - j)
  }
}

print(matrix_result)

```

2. Print the string "*" using for() function. The output should be the same as shown in Figure 2.
```{r}
for (i in 1:5) {
  for (j in 1:i) {
    cat('"*"\t')
  }
  cat("\n\n")
}

```

3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
fib_sequence <- function() {
  start <- as.integer(readline(prompt = "Enter starting number: "))
  
  if (start <= 0) {
    print("Please enter a positive number")
    return()
  }
  
  a <- 0
  b <- 1
  count <- 0
  result <- c()
  
  repeat {
    if (count >= start) {
      if (b > 500) {
        break
      }
      result <- c(result, b)
    }
    
    temp <- a + b
    a <- b
    b <- temp
    count <- count + 1
  }
  print(result)
}

```


4. Import the dataset as shown in Figure 1 you have created previously.
a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result
```{r}

library(readxl)

# Import the Excel file
data <- read_excel("import_march.xlsx")

head(data)

```

b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.
```{r}
male_data <- subset(data, Gender == "Male")
female_data <- subset(data, Gender == "Female")

male_count <- nrow(male_data)
female_count <- nrow(female_data)

cat("Number of Male observations:", male_count, "\n")
cat("Number of Female observations:", female_count, "\n")
```
c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.

```{r}
gender_counts <- table(data$Gender)

# Plot barplot
barplot(
  gender_counts,
  main = "Number of Male and Female Observations",
  xlab = "Gender",
  ylab = "Count",
  col = c("lightblue", "pink"),   # Male = lightblue, Female = pink
  legend.text = TRUE,
  args.legend = list(x = "topright")
)

```

5. The monthly income of Dela Cruz family was spent on the following:
a. Create a piechart that will include labels in percentage.Add some colors and title of
the chart. Write the R scripts and show its output.
```{r}
expenses <- c(60, 10, 5, 25)
categories <- c("Food", "Electricity", "Savings", "Miscellaneous")

percent <- round(expenses / sum(expenses) * 100)
labels <- paste(categories, percent, "%")

pie(expenses,
    labels = labels,
    col = c("orange", "skyblue", "lightgreen", "pink"),
    main = "Monthly Income Distribution of Dela Cruz Family")

```

6. Use the iris dataset.
a.Check for the structure of the dataset using the str() function. Describe what you have seen in the output.
```{r}
data(iris)
str(iris)
```
b. Create an R object that will contain the mean of the sepal.length,sepal.width,petal.length,and petal.width. What is the R script and its result?
```{r}
iris_means <- data.frame(
  Sepal.Length = mean(iris$Sepal.Length),
  Sepal.Width = mean(iris$Sepal.Width),
  Petal.Length = mean(iris$Petal.Length),
  Petal.Width = mean(iris$Petal.Width)
)
print(iris_means)

```

c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write the R script and its result.
```{r}
species_counts <- table(iris$Species)
pie(species_counts,
    main = "Iris Species Distribution",
    col = c("red", "blue", "green"),
    labels = paste(names(species_counts), species_counts))
```

d.Subset the species into setosa, versicolor, and virginica. Write the R scripts
and show the last six (6) rows of each species.
```{r}
setosa <- iris[iris$Species == "setosa", ]
versicolor <- iris[iris$Species == "versicolor", ]
virginica <- iris[iris$Species == "virginica", ]

print("Last 6 rows of Setosa:")
print(tail(setosa, 6))

print("Last 6 rows of Versicolor:")
print(tail(versicolor, 6))

print("Last 6 rows of Virginica:")
print(tail(virginica, 6))

iris$Species <- as.factor(iris$Species)
```
e. Create a scatterplot of the sepal.length and sepal.width using the different
species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal
width and length, labels for the x and y axis, the pch symbol and colors should be based on
the species.
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width,
     main = "Iris Dataset",
     sub = "Sepal width and length",
     xlab = "Sepal Length",
     ylab = "Sepal Width",
     pch = c(16, 17, 18)[as.numeric(iris$Species)],
     col = c("red", "blue", "green")[as.numeric(iris$Species)])

legend("topright",
       legend = levels(iris$Species),
       pch = c(16, 17, 18),
       col = c("red", "blue", "green"))
```
f. Interpret the result.
```{r}
"The scatterplot shows that the three kinds of Iris flowers have different sepal sizes. The red dots (Setosa) have shorter and wider sepals, so they stay together in one group. The blue dots (Versicolor) and green dots (Virginica) look a bit the same, but the green ones usually have longer sepals. This means we can tell Setosa apart easily, but Versicolor and Virginica are a little harder to tell apart."
```

7. Import the alexa-file.xlsx. Check on the variations. Notice that there are ex-tra whitespaces among black variants (Black Dot, Black Plus, Black Show, BlackSpot). Also on the white variants (White Dot, White Plus, White Show, White Spot).

a. Rename the white and black variants by using gsub() function.
```{r}
library(readxl)
library(dplyr)

alexa <- read_xlsx("alexa_file.xlsx")


alexa$variation <- trimws(alexa$variation)   # remove leading/trailing spaces
alexa$variation <- gsub("\\s+", " ", alexa$variation) 
alexa$variation <- trimws(alexa$variation)
alexa$variation <- tools::toTitleCase(tolower(alexa$variation))

unique(alexa$variation)

head(alexa, 10)

```
b. Get the total number of each variations and save it into another object. Save the
object as variations.RData. Write the R scripts. What is its result?
```{r}
variations <- alexa %>%
  count(variation) %>%
  filter(!is.na(n) & n > 0)

save(variations, file = "variations.RData")
print(variations$variation)

```

c.From the variations.RData, create a barplot(). Complete the details of the chart which include the title, color, labels of each bar.
```{r}
if(nrow(variations) > 0){
  barplot(variations$n,
          names.arg = variations$variation,
          main = "All Alexa Variations",
          xlab = "Variation",
          ylab = "Count",
          col = rainbow(nrow(variations)),
          las = 2,
          cex.names = 0.7)
} else {
  plot.new()
  title("No variations to display")
}
```

d. Create a barplot() for the black and white variations. Plot it in 1 frame, side b  y side. Complete the details of the chart.
```{r}
black_vars <- variations %>% 
  filter(grepl("^Black", variation) & !is.na(n) & n > 0)

white_vars <- variations %>% 
  filter(grepl("^White", variation) & !is.na(n) & n > 0)

# Plot side by side
par(mfrow = c(1, 2))

# Black Variations
if(nrow(black_vars) > 0){
  barplot(
    black_vars$n,
    names.arg = black_vars$variation,
    main = "Black Variations",
    xlab = "Variation",
    ylab = "Count",
    col = "black",
    las = 2,
    cex.names = 0.7
  )
} else {
  plot.new()
  title("No Black Variations")
}

# White Variations
if(nrow(white_vars) > 0){
  barplot(
    white_vars$n,
    names.arg = white_vars$variation,
    main = "White Variations",
    xlab = "Variation",
    ylab = "Count",
    col = "gray90",
    las = 2,
    cex.names = 0.7
  )
} else {
  plot.new()
  title("No White Variations")
}

# Reset layout
par(mfrow = c(1, 1))
```


