---
title: 'Work_Sheet #2'
author: "Syndric James Z. Junsay"
date: "2025-09-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1 Create a vector using : operator
a. Sequence from -5 to 5. Write the R code and its output. Describe its output.
```{r setup, include=FALSE}
y <- c(-5:5)
print(x)
```
The output generates a sequence of integers starting from -5 and increasing by 1 up to 5.

b. x <- 1:7. What will be the value of x?
```{r setup, include=FALSE}
x <- c(1:7)
print(x)
```
The value of x is:
[1] 1 2 3 4 5 6 7
## Exercise 2 Create a vector using seq() function
a. seq(1, 3, by=0.2) # specify step size
```{r setup, include=FALSE}
vector <- seq(1, 3, by=0.2)
print (vector)
```
output: [1] 1.0 1.2 1.4 1.6 1.8 2.0 2.2 2.4 2.6 2.8 3.0
The output generates a numeric sequence starting at 1 and ending at 3, incrementing by 0.2, resulting in the values 1.0, 1.2, 1.4, up to 3.0.

## exercise 3
A factory has a census of its workers. There are 50 workers in total. The following
list shows their ages: 34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18.

a. Access 3rd element, what is the value?
```{r setup, include=FALSE}
ages <- c(4, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
thirdage <- ages[3]
print (thirdage)
```
the value is [1] 22

b. Access 2nd and 4th element, what are the values?
```{r setup, include=FALSE}
ages <- c(4, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
second_and_fourth <- ages[c(2,4)]
print (second_and_fourth)
```
the values of the 2nd & 4th elements are [1] 28 36

c. Access all but the 1st element is not included. Write the R code and its output.
```{r setup, include=FALSE}
ages <- c(4, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
all_but_one <- ages[c(2:50)]
print (all_but_one)
```
Rcode:
ages <- c(4, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
all_but_one <- ages[c(2:50)]
print (all_but_one)

Output:[1] 28 22 36 27 18 52 39 42 29 35 31 27 22 37 34 19 20 57 49 50 37 46 25 17 37 43 53
[28] 41 51 35 24 33 41 53 40 18 44 38 41 48 27 39 19 30 61 54 58 26 18
> 

