---
title: 'Work_Sheet #2'
author: "Syndric James Z. Junsay"
date: "2025-09-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1 Create a vector using : operator
#a. Sequence from -5 to 5. Write the R code and its output. Describe its output.
```{r setup, include=FALSE}
y <- c(-5:5)
print(x)
```
#The output generates a sequence of integers starting from -5 and increasing by 1 up to 5.

#b. x <- 1:7. What will be the value of x?
```{r setup, include=FALSE}
x <- c(1:7)
print(x)
```
The value of x is:
[1] 1 2 3 4 5 6 7
## Exercise 2 Create a vector using seq() function
#a. seq(1, 3, by=0.2) # specify step size
```{r setup, include=FALSE}
vector <- seq(1, 3, by=0.2)
print (vector)
```
output: [1] 1.0 1.2 1.4 1.6 1.8 2.0 2.2 2.4 2.6 2.8 3.0
The output generates a numeric sequence starting at 1 and ending at 3, incrementing by 0.2, resulting in the values 1.0, 1.2, 1.4, up to 3.0.

## exercise 3
A factory has a census of its workers. There are 50 workers in total. The following
list shows their ages: 34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18.

a. Access 3rd element, what is the value?
```{r setup, include=FALSE}
ages <- c(4, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
thirdage <- ages[3]
print (thirdage)
```
the value is [1] 22

b. Access 2nd and 4th element, what are the values?
```{r setup, include=FALSE}
ages <- c(4, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
second_and_fourth <- ages[c(2,4)]
print (second_and_fourth)
```
the values of the 2nd & 4th elements are [1] 28 36

c. Access all but the 1st element is not included. Write the R code and its output.
```{r setup, include=FALSE}
ages <- c(4, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
all_but_one <- ages[c(2:50)]
print (all_but_one)
```
Rcode:
ages <- c(4, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
all_but_one <- ages[c(2:50)]
print (all_but_one)

Output:[1] 28 22 36 27 18 52 39 42 29 35 31 27 22 37 34 19 20 57 49 50 37 46 25 17 37 43 53
[28] 41 51 35 24 33 41 53 40 18 44 38 41 48 27 39 19 30 61 54 58 26 18

## Exercise 4 *Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the vector, names(x).

a. Print the results. Then access x[c("first", "third")].
Describe the output.

```{r setup, include= FALSE}
x <- c(3, 0, 9)
names(x) <- c("first", "second", "third")
x[c("first", "third")]
```
the output prints out the values of the names "first" & "third".
    
b.
  code:
    x <- c(3, 0, 9)
    names(x) <- c("first", "second", "third")
    x[c("first", "third")]
  output:
    first third 
    3     9 
    
## Exercise 5 Create a sequence x from -3:2.

a. Modify 2nd element and change it to 0;
x[2] <- 0
x

Describe the output.
```{r setup, include= FALSE}
x <- c(-3:2)
x[2] <- 0
x
```
the output shows that instead of -2 the value was changed to 0

b.
    code:
      x <- c(-3:2)
      x[2] <- 0
      x
    output:
      [1] -3  0 -1  0  1  2

## Exercise 6 The following data shows the diesel fuel purchased by Mr. Cruz.
    Month| Jan, Feb, March, Apr, May, June
    Price per liter (PhP)| 52.50, 57.25, 60.00, 65.00, 74.25, 54.00
    Purchase–quantity(Liters)| 25, 30, 40, 50, 10, 45

a. a. Create a data frame for month, price per liter (php) and purchase-quantity (liter). Write the codes.

```{r setup, include =FALSE}
month <- c("Jan", "Feb", "March", "Apr", "May", "June")
price <-c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
purchase_Quantity <-c(25, 30, 40, 50, 10, 45)

fuel_purchase <- data.frame(month, price, purchase_Quantity)
fuel_purchase
```
b. What is the average fuel expenditure of Mr. Cruz from Jan to June? Note: Use
weighted.mean(liter, purchase)
```{r setup, include= FALSE}
average_fuel <- weighted.mean(price, purchase_Quantity)
average_fuel
```
## 7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths(in miles) of 141 “major” rivers in North America, as compiled by the US Geological Survey”.

a. Type “rivers” in your R console. Create a vector data with 7
elements, containing the number of elements (length) in rivers,
their sum (sum), mean (mean), median (median), variance (var)

```{r setup, include = FALSE}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers), sd(rivers), min(rivers), max(rivers))
data
```
b. What are the results?
[1]    141.0000  83357.0000    591.1844    425.0000 243908.4086
[6]    493.8708    135.0000   3710.0000

c. Write the code and its outputs.
    code:
      data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers),             var(rivers), sd(rivers), min(rivers), max(rivers)) data
    outputs:
      [1]    141.0000  83357.0000    591.1844    425.0000 243908.4086
      [6]    493.8708    135.0000   3710.0000

## 8 The table below gives the 25 most powerful celebrities and their annual pay as ranked by the editions of Forbes magazine and as listed on the Forbes.com website.

a. Create vectors according to the above table. Write the codes.
```{r setup, include = FALSE}
power_ranking <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10,
                   11, 12, 13, 14, 15, 16, 17, 18, 19, 20,
                   21, 22, 23, 24, 25)

celebrity_name <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods",
                    "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos",
                    "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali",
                    "Paul McCartney", "Elon Musk", "Elton John", "David Letterman",
                    "Phil Mickelson", "J.K. Rowling", "Brad Pitt", "Peter Jackson",
                    "Dr. Phil McGraw", "Jay Leno", "Celine Dion", "Kobe Bryant")

# Pay in million USD
pay <- c(67, 90, 225, 110, 90,
         332, 302, 41, 52, 88,
         55, 44, 55, 40, 233, 34,
         40, 47, 75, 25, 39, 45, 32, 40, 31)
celebrities <- data.frame(power_ranking, celebrity_name, pay)
celebrities
```
b. Modify the power ranking and pay of J.K. Rowling. Change power ranking to 15 and
pay to 90. Write the codes and its output.

```{r setup, include = FALSE}
index <- which(celebrity_name == "J.K. Rowling")

# Modify power ranking and pay
power_ranking[index] <- 15
pay[index] <- 90

celebrity_name[index]
power_ranking[index]
pay[index]

#output
#celebrity_name[index]
#[1] "J.K. Rowling"
# power_ranking[index]
#[1] 15
# pay[index]
#[1] 90
> 
```

c. Interpret the data.

The data shows the top 25 highest-paid or most powerful celebrities according to Forbes. power_ranking indicates their influence, while pay indicates their annual earnings in million USD.

Example interpretations:

Highest-paid celebrity: Steven Spielberg ($332M)
Celebrity with lowest pay: Brad Pitt ($25M)

J.K. Rowling’s new ranking and pay indicate a significant increase in both influence and earnings.

This data set can be used to analyze trends between pay and influence, e.g., do higher pay always mean higher ranking?